Texture Mapping 3D Models of Indoor Environments with Noisy Camera Poses

### 2014-01-04 spie2014/spie2014.tex
final perhaps

For the purposes of texturing, it is convenient to treat each region
as planar. To accomplish this, a plane is fitted to each region. This
plane corresponds to the largest contiguous section of coplanar
triangles, as most regions are characterized by small outcroppings
from a largely planar surface, such as objects on a table or features
protruding from a wall. The 3D surface composing a region is then
projected onto its calculated plane, and the resulting 2D planar
polygon is used as the texturing surface. This 2D planar approximation
provides a simple, yet approximately accurate surface onto which
images can be projected in order to generate textures. However, 3D
surfaces are still retained in order to achieve accurate results when
performing occlusion checks, and of course are used to generate the
final textured model. Each region now has a set of triangles
representing the original model geometry, as well as a planar
approximation. The task now is to generate a texture for each region.

@@

There are many existing approaches to stitching together multiple
images to produce a larger, seamless image \cite{szeliski2006image,
  agarwalapanoramas, wangmultipleviews, coorg1997matching,
  debevechybrid, bernardinimultiplescans}. Generally, parts of images
are matched to each other, by detecting feature points and identifying
matches. Images are then transformed to maximizally align matches,
often by computing homographies between pairs of images, or by
iteratively adjusting camera poses in 1 to 6 degrees of freedom.

@@

Our criteria for image subsampling has the following
stipulations. First, we want a set of images such that their
projections together cover up the entirety of our target surface, so
as to avoid holes in the final texture. Second, we desire our final
texture to have high resolution throughout; thus every location on the
target surface should have at least one image that contains high
resolution imagery for it. A straightforward way to accomplish these
goals is to discretize the target surface into small square tiles, and
for each tile, select the image that has the best viewing angle and
resolution for texturing that tile. In order to select the image that
can best texture a tile $t$, we first gather a list of candidate
images that contain all four of its corners; we can rapidly check this
by projecting $t$ into each image using the $P_i$ camera
matrices. Furthermore, each candidate image must have been taken at a
time when its camera had a clear line-of-sight to $t$, which can be
determined using standard ray-polygon intersection tests between the
camera location, $t$, and every other surface, or via the k-d tree
from Section \ref{sec:geometryPartitioning} if present
\cite{rayintersection}.

@@

In order to correctly texture surfaces, it is important to detect and
remove parts of image projections containing texture for occluding
surfaces. For instance, in Figure \ref{fig:occlusion}(a), an image
used to texture the orange target surface also contains part of a gray
occluding surface. We remove this incorrect texture by performing
ray-polygon intersection tests between the camera location and every
surface in our model except the target surface
\cite{rayintersection}. If using planar approximations for regions, as
explained in Section \ref{sec:geometryPartitioning}, the k-d tree is
used instead. These intersection tests are performed repeatedly in a
recursive grid, until all occlusions are detected and corresponding
occluded texture is removed. Occlusion checking works entirely with
geometry, so by ensuring that images match geometry using
\ref{sec:geometryAlignment}'s alignment procedure, texture belonging
to other surfaces is accurately removed, as seen in Figure
\ref{fig:occlusion}(b) vs. \ref{fig:occlusion}(c).

@@

To find the subset of images that covers the plane with minimal cost,
we form a shortest path problem. This can be done because the
wide-angle cameras have full floor-to-ceiling coverage of walls, and
the backpack operator logically only moves horizontally. We thus are
only concerned with lateral coverage when working with wall
surfaces. We can therefore construct a horizontal Directed Acyclic
Graph (DAG) from the images, with edge costs defined by the SSD cost
function, and solve a simple shortest path problem to find an optimal
subset of images with regard to the SSD cost function \cite{dijkstra}.

@@

When texturing our models, we apply the shortest path method on
surfaces where images are taken at optimal angles and provide full
coverage along one dimension. Floors, ceilings, and smaller complex
objects such as furniture, given their many images taken at oblique
angles, are textured using the tile caching method of Section
\ref{sec:mappingWithCaching}.

@@

Similar to the image position refinement procedure in Section
\ref{sec:robustSIFTFeatureMatching}, exposure compensation is
performed simulatenously across all images present in a single
region. To begin, a relative gain $R_{ij}$ is computed for each pair
of overlapping images $I_i$ and $I_j$, which have a set of pixels
common to both images, labeled as $P_{ij}$. The relative gain $R_{ij}$
is then obtained by calculating the scaling factor between the average
intensity of $P_{ij}$ in $I_i$ and the average intensity of $P_{ij}$
in $I_j$.

@@

The texture mapping procedure in this paper was tested with three
different geometry-reconstruction methods, a PCA-based plane-fitting
method \cite{sanchez2012point}, a floor plan extrusion method
\cite{turnerfloorplan}, and a voxel-carving mesh generation method
\cite{turnerwatertight}. The first two methods generate very similar
low-resolution models, containing only major walls, ceilings and
floors. The third method generates high-resolution models, and
attempts to reconstruct all scanned objects in the
environment. Examples of individually textured regions are shown in
Figure \ref{fig:indivPlanes}, while fully texture mapped models
generated by the PCA-based, floor plan extrusion, and voxel-carving
approaches, can be seen in Figures \ref{fig:pcaModels},
\ref{fig:fpModels}, and \ref{fig:voxelModels} respectively.

\subsection{Runtime}
As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:indivPlanes}(a) is a single region, and its texture
was generated with 7543 $\times$ 776 pixels, spanning a 40 meter long
wall. Given 41000 input images in the dataset, a 2.8 GHz dual-core
machine takes under 10 seconds to choose 36 candidate images and
perform the full texturing process. The tile-caching approach takes a
similar amount of time as well, with the much larger floor textures in
Figure \ref{fig:indivPlanes}(c) taking 5-10 minutes to generate.
While not real-time, our process is capable of generating fast updates
after changes in various parameters or modifications to input data,
and if integrated directly into a 3D modeling system, could provide
quick visual feedback as data is collected.

@@

In this paper, we have developed an approach to texture mapping models
of indoor environments with noisy camera localization data. We are
able to refine image locations based on geometry references and
feature matching, and robustly handle outliers. Using the tile-based
mapping approach, we can texture both large planar features as well as
smaller, more complex surfaces. We also implemented a shortest path
texturing method that produces seamless textures on planes where
multiple head-on images are available. Both of these approaches are
highly modular, and easily tunable for similar systems across multiple
environments.

Our method is likely to fail however in scenarios where 3D error is
large. A logical progression of our approach to resolve camera error
in 3D is to perform matching between image lines and geometry in 3D,
which can be done reasonably efficiently \cite{linebased,
  rectangularstructures}. Using linear features in addition to SIFT
features is also likely to result in improved matches, as indoor
scenes often have long, unbroken lines spanning multiple images
\cite{linearposeestimation}. Finally, the blending procedure is quite
basic, and applying more sophisticated methods of blending as well as
image boundary selection would benefit the final visual quality, and
more robustly handle motion-based or parallax errors.

>>>>

For the purposes of texturing, it is convenient to treat each region
as planar, and fit a plane to each region. This plane corresponds to
the largest contiguous section of coplanar triangles in the region, as
most regions are characterized by small outcroppings from a largely
planar surface, such as objects on a table or features protruding from
a wall. The 3D surface composing a region is then projected onto its
calculated plane, and the resulting 2D planar polygon is used as the
texturing surface. This 2D planar approximation provides a simple, yet
approximately accurate surface onto which images can be projected in
order to generate textures. However, 3D surfaces are still retained in
order to achieve accurate results when performing occlusion checks,
and of course are used to project the final textures upon. Each region
now has a set of triangles representing the original model geometry,
as well as a planar approximation. The task now is to generate a
texture for each region.

@@

There are many existing approaches to stitching together multiple
images to produce a larger, seamless image \cite{szeliski2006image,
  agarwalapanoramas, wangmultipleviews, coorg1997matching,
  debevechybrid, bernardinimultiplescans}. Generally, parts of images
are matched to each other, by detecting feature points and identifying
matches. Images are then transformed to maximally align matches,
often by computing homographies between pairs of images, or by
iteratively adjusting camera poses in 1 to 6 degrees of freedom.

@@

Our criteria for image subsampling has the following
stipulations. First, we want a set of images such that their
projections together cover up the entirety of our target surface, so
as to avoid holes in the final texture. Second, we desire our final
texture to have high resolution throughout; thus every location on the
target surface should have at least one image that contains high
resolution imagery for it. A straightforward way to accomplish these
goals is to discretize the target surface into small square tiles, and
for each tile, select the image that has the best viewing angle and
resolution for texturing that tile. In order to select the image that
can best texture a tile $t$, we first gather a list of candidate
images that contain all four of its corners; we can rapidly check this
by projecting $t$ into each image using the $P_i$ camera
matrices. Furthermore, each candidate image must have been taken at a
time when its camera had a clear line-of-sight to $t$, which can be
determined using standard ray-polygon intersection tests between the
camera location, $t$, and the rest of the model's geometry.
\cite{rayintersection}.

@@

In order to correctly texture surfaces, it is important to detect and
remove parts of image projections containing texture for occluding
surfaces. For instance, in Figure \ref{fig:occlusion}(a), an image
used to texture the orange target surface also contains part of a gray
occluding surface. We remove this incorrect texture by performing
ray-polygon intersection tests between the camera location and every
surface in our model except the target surface
\cite{rayintersection}. These intersection tests are performed repeatedly in a
recursive grid, until all occlusions are detected and corresponding
occluded texture is removed. Occlusion checking works entirely with
geometry, so by ensuring that images match geometry using
\ref{sec:geometryAlignment}'s alignment procedure, texture belonging
to other surfaces is accurately removed, as seen in Figure
\ref{fig:occlusion}(b) vs. \ref{fig:occlusion}(c).

@@

To find the subset of images that covers the plane with minimal cost,
we form a shortest path problem. This can be done because the
wide-angle cameras have full floor-to-ceiling coverage of walls, and
the backpack operator logically only moves horizontally. We thus are
only concerned with lateral coverage when working with wall
surfaces. We can therefore construct a horizontal Directed Acyclic
Graph (DAG) from the images, with edge costs defined by the SSD cost
function, and solve a simple shortest path problem to find an optimal
subset of images with which to texture the surface \cite{dijkstra}.

@@

When texturing our models, we apply the shortest path method on
surfaces where images are taken at optimal angles and provide full
coverage along one dimension. This generally includes all walls and
most other vertical surfaces. Floors, ceilings, and smaller complex
objects such as furniture, given their many images taken at oblique
angles, are textured using the tile caching method of Section
\ref{sec:mappingWithCaching}.

@@

Similar to the image position refinement procedure in Section
\ref{sec:robustSIFTFeatureMatching}, exposure compensation is
performed simultaneously across all images present in a single
region. To begin, a relative gain $R_{ij}$ is computed for each pair
of overlapping images $I_i$ and $I_j$, which have a set of pixels
common to both images, labeled as $P_{ij}$. The relative gain $R_{ij}$
is then obtained by calculating the scaling factor between the average
intensity of $P_{ij}$ in $I_i$ and the average intensity of $P_{ij}$
in $I_j$.

@@

The texture mapping procedure in this paper was tested with three
different geometry-reconstruction methods, a PCA-based plane-fitting
method \cite{sanchez2012point}, a floor plan extrusion method
\cite{turnerfloorplan}, and a voxel-carving mesh generation method
\cite{turnerwatertight}. The first two methods generate low-resolution
models containing only major walls, ceilings and floors. The third
method generates high-resolution models, and attempts to reconstruct
all scanned objects in the environment. Examples of individually
textured regions are shown in Figure \ref{fig:indivPlanes}, while
fully texture mapped models generated by the PCA-based, floor plan
extrusion, and voxel-carving approaches, can be seen in Figures
\ref{fig:pcaModels}, \ref{fig:fpModels}, and \ref{fig:voxelModels}
respectively.

\subsection{Runtime}
As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:indivPlanes}(a) is a single region, and its texture
was generated with 7543 $\times$ 776 pixels, spanning a 40 meter long
wall. Given 41000 input images in the dataset, a 2.8 GHz dual-core
machine takes under 30 seconds to choose 36 candidate images and
perform the full texturing process. The tile-caching approach takes a
similar amount of time as well, with the larger floor textures in
Figure \ref{fig:indivPlanes}(c) taking 5-10 minutes to generate.
While not real-time, our process is capable of generating fast updates
after changes in various parameters or modifications to input data,
and if integrated directly into a 3D modeling system, could provide
quick visual feedback as data is collected.

@@

In this paper, we have developed a method for texture mapping models
of indoor environments with noisy camera localization data. We are
able to refine image locations based on geometry references as well as
feature matching, while robustly handling environments with repeating
or sparse details. Using the tile-based mapping approach, we can
texture both large planar features as well as smaller, more complex
surfaces. We also implemented a shortest path texturing method that
produces seamless textures on planes where multiple head-on images are
available. Both of these approaches are highly modular, and easily
tunable for similar systems across multiple environments.

Our method is likely to fail however in scenarios where 3D error is
large. A logical progression of our approach to resolve camera error
in 3D is to perform matching between image lines and geometry in 3D,
which can be done reasonably efficiently \cite{linebased,
  rectangularstructures}. Using linear features in addition to SIFT
features is also likely to result in improved matches, as indoor
scenes often have long, unbroken lines spanning multiple images
\cite{linearposeestimation}. Finally, the blending procedure is quite
basic, and applying more sophisticated methods of blending as well as
image boundary selection would benefit texture visual quality, and
more robustly handle motion-based or parallax errors.
